---
title: "Why do Individuals Support Automatic Registration?"
author: "Mena Solomon"
description: "A report exploring the effect of support for progressive voting policy, like abolishing voter ID laws, on an individual's perspective on automatic registration."
output:
  distill::distill_article:
    self_contained: false
---



Introduction
----

With the 2024 general election rapidly approaching, issues of low voter turnout and mass disenfranchisement continue to plague this country. Policies which seek to reach the quarter of Americans who are not currently registered to vote continue to garner the attention of both politicians and the media. In particular, the policy of automatic registration, which would replace the opt-in system of voter registration with an opt-out system and automatically register everyone who turns 18, has already been successfully implemented in over 20 states. Indeed, a report from the New Brennan Center found that "AVR markedly increases the number of voters being registered [with] increases in the number of registrants ranging from 9 to 94 percent." If automatic registration could be the key to increasing voter registration and thus turnout in the United States, why do some individuals still refuse to support this policy? This report seeks to understand why certain individuals support automatic voter registration, and to what extent an individual’s perspective on automatic registration is merely a product of their support of other progressive voting policies.

I hypothesize that perspective on automatic registration will correlate with perspective on voter ID laws, wherein individuals who oppose voter ID laws, a policy with a longstanding history of voter discrimination, will support automatic registration, a policy which would increase access to voting. Furthermore, I expect this relationship to be in large part explained first by party membership, as democrats typically support progressive voting policies, and second by racial identity, as marginalized racial groups are historically the most affected by voter disenfranchisement.  

The Data
----

The data for this report comes from the Pew Research Center's American Trends Panel, a survey which is sent to a random sampling of 10,000 American adults who serve as a microcosm of America today. The Pew Center makes a conscious effort to minimize the effect of external variables on the data they collect, thus generating trends data that is representative of American opinion. This report analyzes data on the opinion's of respondents on various forms of voter registration — specifically voter ID laws and automatic registration — in the months after the 2020 general election. This data comes from Wave 87 of the American Trends Panel which was conducted in April 2021, three months after the Inauguration of President Biden, and consists of responses from half of the panelists (5,109 respondents).

Since the American Trends Panel contains data on a multitude of issues, it was important to filter, rename, and refine the data to address the research question stated above. The independent variable in this research is perspective on voter ID laws, reported in the variable VTPRIORITY_j_W87 where  1 = Strongly favor, 2 = Somewhat favor, 3 = Somewhat oppose, and 4 = Strongly oppose. In the analysis below, this variable was renamed to voteridlaws. Furthermore, to simplify this variable for the data visualization, a new variable was created entitled voteridlabel. For this variable, those who responded with a one or two were grouped into a category of supporting voter ID laws and those who responded with a 3 or 4 were grouped as being against voter ID laws. The dependent variable is perspective on automatic registration, reported in the variable VTPRIORITY_a_W87 which operates on the same scale as the voter ID law variable. This variable was similarly renamed to be automaticregistration, then further split into two categories of support and oppose under the variable automaticreglabel. 

Furthermore, this report also measures the influence of other explanatory (independent) variables such as race, age, gender, and political party. These variables were all imported into the original data under the names F_AGECAT, F_GENDER, F_RACETHNMOD, and F_PARTY_FINAL and renamed as age, gender, race, and party respectively. In the data set, age is broken up into four categories 18-29, 30-49, 50-64, and 65+. Gender was measured as 1 = respondent identifies as a man, 2 = respondent identifies as a women, 3 = respondent identifies in some other way. Race is broken up into five categories where 1 = White, non-Hispanic, 2 = Black, non-Hispanic, 3 = Hispanic, 4 = other, and 5 = Asian, non-Hispanic. Finally, party was broken down into 1 = Republican, 2 = Democrat, 3 = Independent, and 4 = Something else

For all of the variables stated above, the option of non-response was filtered out. For most of the variables, non-response was represented by a result of 99. For party and race, however, an answer of four represented an answer of other, which, for the sake of this analysis, was removed. Furthermore, the answer of Independent for party affiliation was removed for the sake of simplicity. These alterations filtered the data from 5,109 responses to 2,902 responses, it is those individuals whose data is analyzed below.

```{r}
library(tidyverse)
library(haven)
library(infer)
data <-  read_sav("data/ATP W87.sav")

data <- data |>
  select(QKEY, F_AGECAT, F_GENDER, F_RACETHNMOD, F_PARTY_FINAL, VTPRIORITY_a_W87, VTPRIORITY_j_W87)|>
  rename(automaticregistration = VTPRIORITY_a_W87,
         voteridlaws = VTPRIORITY_j_W87,
         party = F_PARTY_FINAL,
         race = F_RACETHNMOD,
         gender = F_GENDER,
         age = F_AGECAT) |>
  filter(party != 4,
         party != 3,
         party != 99, 
         voteridlaws != 99,
         automaticregistration != 99,
         race != 4,
         race != 99,
         gender != 99,
         age != 99)|>
  mutate(
    partylabel = case_when(
      party == 1 ~ "Republican",
      party == 2 ~ "Democrat"),
    racelabel = case_when(
      race == 1 ~ "White, non-Hispanic",
      race == 2 ~ "Black, non-Hispanic",
      race == 3 ~ "Hispanic",
      race == 5 ~ "Asian, non-Hispanic"),
    voteridlabel = case_when(
      voteridlaws <= 2 ~ "Supports Voter ID Laws",
      voteridlaws > 2 ~ "Against Voter ID Laws"),
    automaticreglabel = case_when(
      automaticregistration <= 2 ~ "Supports Automatic Registration",
      automaticregistration > 2 ~ "Against Automatic Registration"))


```

Data Visualizations
---

The figure below illustrates the dependent variable: respondent's perspectives on automatic registration.

```{r}
my_colors <- c("Strongly Supports" = "cadetblue1", 
               "Somewhat Supports" = "cadetblue3", 
               "Strongly Against" = "indianred1", 
               "Somewhat Against" = "indianred3")

autoregplot <- data |>
  mutate(automaticregdetail = case_when(automaticregistration == 1 ~ "Strongly Supports",
                                        automaticregistration == 2 ~ "Somewhat Supports",
                                        automaticregistration == 3 ~ "Strongly Against",
                                        automaticregistration == 4 ~ "Somewhat Against")) |>
  group_by(automaticreglabel) |>
  ggplot(mapping = aes(x = automaticreglabel, fill = factor(automaticregdetail))) +
  geom_bar() +
  scale_fill_manual(values = my_colors) +
  labs(x = "Opinion on Automatic Registration", 
       y = "Number of Respondents", 
       title = "How do Respondents Feel About Automatic Registration",
       fill = "Degree of Support")
autoregplot
```
This figure displays that nearly 1,500 respondents strongly support automatic registration, with over 2,000 supporting the policy to some degree. Indeed, only 750 respondents reported being against the policy.

The next figure illustrates the independent variable: respondent's perspectives on voter ID laws.

```{r}
voteridplot <- data |>
  mutate(voteriddetail = case_when(voteridlaws == 1 ~ "Strongly Supports",
                                        voteridlaws == 2 ~ "Somewhat Supports",
                                        voteridlaws == 3 ~ "Strongly Against",
                                        voteridlaws == 4 ~ "Somewhat Against")) |>
  group_by(voteridlabel) |>
  ggplot(mapping = aes(x = voteridlabel, fill = factor(voteriddetail))) +
  geom_bar() +
  scale_fill_manual(values = my_colors) +
  labs(x = "Opinion on Voter ID Laws", 
       y = "Number of Respondents", 
       title = "How do Respondents Feel About Voter ID Laws",
       fill = "Degree of Support")
voteridplot
```
Similar to automatic registration, there appears to be more support that opposition for voter ID laws. That said, there are fewer individuals — under 1500 — who support these laws strongly and nearly 500 respondents strongly oppose this policy.

Results
---

Bootstrapped Difference in Means Calculation
---

To understand how a respondent's perspective on voter ID laws relates to their perspective on automatic registration, I performed a difference in means calculation wherein the average feeling toward automatic registration was taken for both the group who supports voter ID laws and those who don't, then subtracted from each other. In an attempt to limit the interference of outside variables, bootstrap replications were used, generating 1,000 random sample groups. 


```{r}
ate_boots <- data |>
  rep_slice_sample(prop = 1, reps = 1000, replace = TRUE) |>
  group_by(replicate, voteridlabel) |>
  summarize(automaticregistrationmean = mean(automaticregistration)) |>
  pivot_wider(names_from = voteridlabel, values_from = automaticregistrationmean) |>
  mutate(ATE = `Supports Voter ID Laws` - `Against Voter ID Laws`)

ate_ci_95 <- ate_boots |>
select(replicate, ATE) |>
get_confidence_interval(level = 0.95, type = "percentile")

ate_boots |>
  ggplot(mapping = aes(x = ATE))+
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.01) +
  geom_vline(xintercept = unlist(ate_ci_95), color = "red") +
  labs(x = "Difference in Means", 
       y = "Density", 
       subtitle = "1,000 Bootstrap Replications of the Estimated Affect of Supporting Voter ID Laws on Feelings Toward Automatic Registration",
       title = "How Perspective on Voter ID Laws Affects Views on Automatic Registration")

```
The bootstrap replications display the number of samples which result in a particular difference in means. For the graph of views on automatic registration, the average treatment effect follows a normal distribution with the center located around 0.57, meaning on average respondents who support voter ID laws were less in favor of automatic registration than those who did not by about 0.57 points. Furthermore, the 95% confidence interval calculated above  does not include zero. Thus, at least at the 95% confidence interval, it is possible to discount the null hypothesis — support for voter ID laws bears no effect on a respondent's views on automatic registration.

Multiple Regression
---

To further explore this relationship, I conducted a regression which not only analyzed the impact of my primary independent variable but also my other explanatory variables. This regression analyzes the effect of respondents' perspective on voter ID laws, race, party, gender, and age on their views about automatic registration.

```{r}
library(infer)
library(modelsummary)
data2 <- data |>
  filter(race != 3,
         race != 5,
         gender !=3) |>
  mutate(voteridbinary = if_else(voteridlaws < 2, 1, 0))

fit_ate_voterid <- 
  lm(automaticregistration ~ voteridbinary + race + party + gender + age, 
     data = data2)

varnames <- c("(Intercept)" = "Intercept",
               "voteridbinary" = "Perspective on Voter ID Laws",
               "race" = "Race",
               "party" = "Political Party Affiliation",
               "gender" = "Gender",
               "age" = "Age")

modelsummary::modelsummary(fit_ate_voterid,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_map = varnames)  
```
This regression indicates that relative to those who oppose voter ID laws, an individual who supports voter ID laws rates themselves 0.148 points higher on their feeling towards automatic registration: meaning they oppose the policy by 0.148 points more than their peers who do not support voter ID laws. The p-value of 0.005 indicates that this result is highly significant. Furthermore, this p-value asserts that we are not only able to reject the null hypothesis of no effect at the 0.05 level (found above) but at the 0.01 level as well, as 0.005 is less than 0.01.

The regression also analyzes the effect of other independent, and possibly explanatory, variables. To evaluate if the effect of these other variables is statistically significant, it is important to first look at their p-values. Only age and party have a p-value under 0.05, the value which indicates statistical significance. This data indicates that neither race nor gender have a significant impact on a respondent's view toward automatic registration. Furthermore, their p-values indicate that we can reject the null hypothesis — that, analyzed separately, age and party have no effect on a respondent's opinion on automatic registration — at the 0.01 level.

In examining age and party more closely, the coefficient for age is found to be 0.066, indicating that as an individual's age increases, they tend to oppose automatic registration by 0.066 more points, relative to younger respondents. The p-value of 0.004 indicates that the coefficient found in the regression is highly significant. Furthermore, the coefficient for party is -1.419, indicating that Democrats tend to favor automatic registration by 1.419 points more than republicans. With a p-value less than 0.001, this result appears to be highly statistically significant. Indeed, the coefficient for party as well as its corresponding p-value indicate that it is party, not age or perspective on voter ID laws, which serves as the largest determinant of a respondent's perspective on automatic registration.

Summary Graph 
---

This final graph summarizes the relationship between party, the statistically significant explanatory variable with the largest coefficient, and views on automatic registration.

```{r}
my_colors2 <- c("Democrat" = "cadetblue1",
                "Republican" = "indianred1")

automatic_registration_plot_party <- data |>
  group_by(party) |>
  ggplot(aes(x = factor(automaticreglabel), fill = factor(partylabel))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = my_colors2) +
  labs(x = "Opinion on Automatic Registration", 
       y = "Number of Respondents", 
       title = "How Party Membership Affects Respondents Views on Automatic Registration",
       fill = "Party")

automatic_registration_plot_party
```
This graph indicates a clear relationship between party membership and support for automatic registration, wherein around 1750 members of the democratic party report supporting automatic registration to some degree. Furthermore, while around 500 members of the republican party oppose automatic registration, under 250 democrats feel similarly. 

Conclusion 
---

The initial hypothesis stated that perspective on automatic registration will correlate with perspective on voter ID laws wherein individuals who oppose voter ID laws, a policy with a longstanding history of voter discrimination, will support automatic registration, a policy which would increase access to voting. Both the bootstrapped difference in means and the regression model indicate that those who support voter ID laws tend to oppose automatic registration. That said, while the null hypothesis of no effect was rejected at the 0.01 level, the effect was found to be quite small. The next part of the hypothesis stated that I expected this relationship to be in large part explained first by party membership, as democrats typically support progressive voting policies, and second by racial identity, as marginalized racial groups are historically the most affected by voter disenfranchisement. Surprisingly, the regression model found that the relationship between race and perspective on automatic registration was not statistically significant. Two other independent variables — age and party membership — were, however, statistically significant, and for both, the null hypothesis of no effect could be rejected. Indeed, an increase in age appeared to correlate with a small increase in opposition toward automatic registration. Furthermore, the coefficient for party membership was the largest, indicating that Democrats tended to favor automatic registration by 1.419 points more than republican respondents. 

These results beget the conclusion that passing automatic registration policy is not seen as the same as abolishing voter ID laws in the United States. Indeed, while those who support voter id laws correlate somewhat with those who support automatic registration, the explanatory variable with the largest impact on perspective is party membership. Despite this, there is not enough data in this study to determine any sort of causal relationship, it merely asserts that there is a definite correlation. To conduct a more comprehensive analysis of the various influences that determine an individual's perspective on voter registration, it is important to look at many policies across time. This would require conducting a longitudinal study to analyze how perspectives on automatic registration have changed in response to various changes in both policy and discourse. Furthermore, by asking more in-depth questions, it would be possible to perform in-depth party analysis as well, assessing if particular sects of the democratic party — be it identity groups, regional groups, or others — favor certain policies over others. Through detailed analysis of the mechanisms by which individuals come to support certain voter registration policies, like automatic registration, it will be possible to formulate politically feasible policies that will truly expand voting access and combat disenfranchisement in the United States.

Works Cited
---

Keeter, Scott. "Growing and Improving Pew Research Center's American Trends Panel." Pew Research 
     Center. Last modified February 27, 2019. Accessed December 13, 2023. 
     https://www.pewresearch.org/methods/2019/02/27/ 
     growing-and-improving-pew-research-centers-american-trends-panel/. 
 
Morris, Kevin, and Peter Dunphy. AVR Impact on State Voter Registration. April 11, 2019. Accessed 
     December 13, 2023. https://www.brennancenter.org/our-work/research-reports/ 
     avr-impact-state-voter-registration. 
 
Republicans and Democrats Move Further Apart in Views of Voting Access. April 22, 2021. Accessed 
     December 13, 2023. https://www.pewresearch.org/politics/2021/04/22/ 
     republicans-and-democrats-move-further-apart-in-views-of-voting-access/. 
 